---
title: "Introduction to Probability"
author: "Madhuka Gamage and Rajinda Wickrama"
date: "April 14, 2025"
site: bookdown::bookdown_site
output:
  bookdown::gitbook:
    config:
      toc:
        collapse: section
        before: |
          <li><strong>Contents</strong></li>
  bookdown::pdf_book:
    latex_engine: xelatex
    citation_package: natbib
    keep_tex: true
    includes:
      in_header: preamble.tex
    toc: true
    number_sections: true
    fig_caption: true
documentclass: krantzNoCorner
fontsize: 12pt
colorlinks: yes
github-repo: https://github.com/openprobabilitytext/introprob2025/
---

# Preface {-}

This online text was developed specifically for students enrolled in MATH350- Probability Theory and Simulation Methods at the University of Delaware. It is designed to accompany our lectures, assignments, and discussions, providing a structured and focused introduction to key concepts in probability.

The book covers topics including fundamentals of probability, counting techniques, conditional probability, independence, random variables, expectations, common univariate distributions, joint distributions, and limit theorems. It is not intended as a comprehensive reference, but rather as a concise, course‑aligned guide to help you develop intuition and problem‑solving skills in probability.

In addition to serving our course, this text is also an excellent resource for MATH367-Actuarial Seminar: Exam P Preparation course. Key concepts and problem types map directly to the Exam P syllabus, and you will find targeted examples and practice questions throughout to reinforce your understanding.

You are encouraged to read actively; work through examples, pause to answer questions, and revisit tricky sections. Probability takes time to absorb, and deeper understanding often comes through repeated engagement with core ideas.

This resource is freely available online and can also be downloaded as a PDF for offline use. The online format allows for updates, so check periodically for revised versions, new examples, or additional exam‑style questions.

We welcome student feedback, which can help improve the book for future readers. We hope you find this a helpful and accessible resource as you build your understanding of probability and prepare for Exam P.


```{r setup, include=FALSE}
knitr::opts_chunk$set(tidy = T, echo = TRUE, comment=NA, message = FALSE, warning = FALSE, fig.align = "center")
HtmlEval <- knitr:::is_html_output()
PdfEval  <- knitr:::is_latex_output()

```

```{r echo = FALSE}
#  Some Switches
QUIZ              <- TRUE
ANIMATION         <- TRUE
SHOW_PDF          <- TRUE
SHOW_EXAMPLE_SOLN <- TRUE
SHOW_EXER_SOLN    <- TRUE
ANIMATIONHOOK     <- 'gifski'


if (PdfEval){
QUIZ              <- FALSE
ANIMATION         <- FALSE
SHOW_PDF          <- FALSE
SHOW_EXER_SOLN    <- FALSE
ANIMATIONHOOK     <- NULL
numAnimation = 1
}


roundCom  <- function(number, digits) {
  rounded_number <- round(number, digits)
  formatted_number <- format(rounded_number, big.mark = ",", nsmall = digits)
  return(formatted_number)
}

CodeFontSmall <- function(A){ '\\scalefont{0.75}'    }
CodeFontLarge <- function(A){ '\\scalefont{1.33333}' }

SolnBegin <- function(A){ if (HtmlEval) {                    knitr::asis_output(paste("*Example Solution.*")) } 
  else { '\\begin{graybox}\\scalefont{0.90}
    \\textbf{Example Solution.}' }  }
SolnEnd <- function(A){ 
   if (PdfEval) { '\\end{graybox}' } }

LObjBegin <- function(A){ if (PdfEval){ '\\begin{graybox}\\scalefont{0.90}' } else
  {knitr::asis_output(paste("***")) } }
LObjEnd <- function(A){  if (PdfEval) { '\\end{graybox}' } else
  {knitr::asis_output(paste("***")) } }

Refer <- function(A){
if (PdfEval) { 'hello' } } #Inside: You can learn more about the R code for this example at the online version of this book.

#Only return argument for glossary function
Gloss <- function(term, section){
  return(term)
}

HideExample <- function(arg1,arg2){
  return("")
}

HideRCode <- function(arg1,arg2){
  return("")
}

HideProofTheory <- function(arg1,arg2){
  return("")
}


```
